|

### 2019-07-31 article/article.Rmd
light editing intro

One of the key determinants of efficient public service provision is a robust accountability system [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent; @o1998horizontal]. In its standard definition, accountability is understood as the process of holding authorities into account for their actions [@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past studies show that accountability ensures that politicians act on behalf of voters [@freire2010ngp; @moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking and corruption [@deininger2005does; @wenar2006accountability] and increases the efficiency of the public sector [@adsera2003you; @bjorkman2009power]. Moreover, greater accountability is associated with higher levels of economic growth, not only because it limits state discretion in the economy, but also because it promotes social cohesion and long-term investments in human capital [@benhabib2010economic; @suebvises2018social; @ponzetto2018social].

However, accountability systems can take many forms. One promising model for democratic oversight is _bottom-up accountability_, in which citizens receive information about the shortcomings of a given project to help them monitor and pressure underperforming public officials [@kosack2014does; @molina2016community; @raffler2018weakness]. Proponents suggest bottom-up accountability is effective because 1) constituents have first-hand information about the outcomes of local policies; 2) citizens have incentives to avoid corruption that directly affects them; 3) policy-makers are particularly sensitive to social punishment from their own communities [@serra2011combining, 570]. In a nutshell, bottom-up accountability offers a potential solution to the principal-agent dilemma in public service by aligning the interests of politicians and bureaucrats with those of the constituency they serve [@barro1973control; @raffler2018weakness, 2].

@@

The _Tá de Pé_[^tdp] cell phone application is an initiative carried out by Transparência Brasil to foster bottom-up accountability in the Brazilian public sector. More specifically, the main goal of the TDP project is to improve the responsiveness in education infrastructure public expenditures. As we mentioned above, TDP incentivises citizens to provide up-to-date information about unfinished schools in their neighbourhoods, and these pictures are then assessed by engineers. In the case the construction is behind schedule, citizens are also provided with tools to contact their representatives quickly and anonymously. The app sends a notification to the mayor's office which has 15 days to answer. If they fail answer by the deadline, a notification goes to the federal government, making it harder to the municipality to access federal funds in the future. The motivation for the intervention is that information delivery causes individuals to closely monitor public works. This, in turn, results in better social outcomes as public agents become more responsive to community demands.

>>>>

One of the key determinants of efficient public service provision is a robust accountability system [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent; @o1998horizontal]. In its standard definition, accountability is understood as the process of holding authorities responsible for their actions [@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past studies show that accountability ensures that politicians act on behalf of voters [@freire2010ngp; @moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking and corruption [@deininger2005does; @wenar2006accountability] and improves the efficiency of the public sector [@adsera2003you; @bjorkman2009power]. Moreover, greater accountability leads to higher economic growth as it limits state discretion in the economy, promotes social cohesion and increases long-term investments in human capital [@benhabib2010economic; @suebvises2018social; @ponzetto2018social].

However, accountability systems can take many forms. One promising model for democratic oversight is _bottom-up monitoring_, in which citizens receive information about the shortcomings of a given project to help them evaluate and pressure underperforming public officials [@kosack2014does; @molina2016community; @raffler2018weakness]. Proponents suggest bottom-up accountability is effective because 1) constituents have first-hand information about the outcomes of local policies; 2) citizens have incentives to avoid corruption that directly affects them; 3) policy-makers are particularly sensitive to social punishment from their own communities [@serra2011combining, 570]. In a nutshell, bottom-up accountability offers a potential solution to the principal-agent dilemma in public service by aligning the interests of politicians and bureaucrats with those of the constituency they serve [@barro1973control; @raffler2018weakness, 2].

@@

The _Tá de Pé_[^tdp] cell phone application is an initiative carried out by Transparência Brasil to foster bottom-up accountability in the Brazilian public sector. More specifically, the main goal of the TDP project is to improve responsiveness in government education expenditures. The TDP app incentivises citizens to provide up-to-date information about unfinished schools in their neighbourhoods, and the data will be assessed by a group of independent specialists. In the case the construction is behind schedule, TDP provides a writing platform whereby citizens can contact their representatives quickly and anonymously. The app then sends a notification to the mayor's office, which has 15 days to reply. If they do not respond to the request, a notification goes to the federal government, making it harder to the municipality to access federal funds in the future. The motivation for the intervention is that information delivery causes individuals to closely monitor public works. This, in turn, results in better social outcomes as public agents become more responsive to community demands.
